---
title: "Capstone Project Quiz 3"
output: html_document
---

Data filepath

```{r}
library(tm)
library(stringr)
library(ngram)
```

```{r}
co_twitter_en <- "/users/petramala/downloads/final/en_US//en_US.twitter.txt"
co_news_en <- "/users/petramala/downloads/final/en_US//en_US.news.txt"
co_blogs_en <- "/users/petramala/downloads/final/en_US/en_US.blogs.txt"
```

Prepare function for tidying data

```{r}
tidyText <- function(file, tidyfile) {
  
  # read in text
  con <- file(file, open="r")
  lines <- readLines(con, skipNul=TRUE)
  close(con)

  lines <- tolower(lines)
  
  # replace words that contain "@", "#", "http://", "https://" 
  # with space (especially for Twitter text)
  lines <- gsub("([^[:space:]]*)(@|#|http://|https://)([^[:space:]]*)", " ", lines)
  
  # split at all ".", ",", brackets and etc.
  lines <- unlist(strsplit(lines, "[.,:;!?(){}<>]|[][]+"))

  # replace all non-alphanumeric characters with a space at the beginning/end of a word.
  lines <- gsub("^[^a-z0-9]+|[^a-z0-9]+$", " ", lines) # at the begining/end of a line
  lines <- gsub("[^a-z0-9]+\\s", " ", lines) # before space
  lines <- gsub("\\s[^a-z0-9]+", " ", lines) # after space
  
  # split a string at spaces then remove the words 
  # that contain any non-alphabetic characters (excpet "-", "'")
  # then paste them together (separate them with spaces)
  lines <- unlist(lapply(lines, function(line){
    words <- unlist(strsplit(line, "\\s+"))
    words <- words[!grepl("[^a-z'-]", words, perl=TRUE)]
    paste(words, collapse=" ")}))
  
  # remove axcess spaces
  #lines <- gsub("\\s+", " ", lines) # remove mutiple spaces
  lines <- str_trim(lines) # remove spaces at the beginning/end of the line

  # drop blank lines
  lines <- lines[nchar(lines)>0]
  
  saveRDS(lines, file = tidyfile) 
}
```

Clean texts

```{r}
tidyText(co_twitter_en, "co_tidy_twitter_en.rds")

tidyText(co_news_en, "co_tidy_news_en.rds")

tidyText(co_blogs_en, "co_tidy_blogs_en.rds")
```

Merge texts

```{r}

df_news <- readRDS(file = "co_tidy_news_en") # 340061 lines

df_blogs <- readRDS(file = "co_tidy_blogs_en") # 4532671 lines

df_twitter <- readRDS(file = "co_tidy_twitter_en") # 5030042 lines

lines <- c(df_news, df_blogs, df_twitter)
#rm(df_news, df_blogs, df_twitter)
```

Get 3-gram

```{r}
# remove lines that contain less than 3 words, or ngram() would throw errors.
lines <- lines[str_count(lines, "\\s+")>1] # reduced 9902774 lines to 7607099 lines

trigram <- ngram(lines, n=3); 
#rm(lines) # 7.619798 mins

df <- get.phrasetable(trigram); 
#rm(trigram) # 3.286831 mins

saveRDS(df, "co_3gram_en.rds") # 211,607 KB
```

The 3-gram dictionary has a long tail.

```{r}
df <- readRDS("co_3gram_en.rds") # 22662982 objects
```

```{r}
plot(df[1:5000,]$freq,
     main='3-Grams Top 5000 Word Frequence',
     ylab="Frequence",
     xlab="Word")
```

```{r}
rbind(head(df, 10), tail(df, 10))
```

Remove 3-grams with freqence = 1 (the long tail)

```{r}
df_notail <- df[df$freq>1,]
saveRDS(df_notail, "co_3gram_notail_en.rds") # 32,995 KB
```

```{r}
df_notail <- readRDS("co_3gram_notail_en.rds")
dim(df_notail)
```

The reduced 3-gram dictionary contains 17.91% original 3-grams, covers 62.45% instances, sizes 15% of the full 3-gram dictionary.

```{r}
c(nrow(df_notail)/nrow(df),
sum(df_notail$freq)/sum(df$freq),
32995/211607)
```

### Quiz 3

Question 1:
For each of the sentence fragments below use your natural language processing algorithm to predict the next word in the sentence.

When you breathe, I want to be the air for you. I'll be there for you, I'd live and I'd

```{r}
rbind(df[grep("^and i'd give", df[,1]),],
      df[grep("^and i'd sleep", df[,1]),],
      df[grep("^and i'd eat", df[,1]),],
      df[grep("^and i'd die", df[,1]),])
```

```{r}
rbind(df_notail[grep("^and i'd give", df_notail[,1]),],
      df_notail[grep("^and i'd sleep", df_notail[,1]),],
      df_notail[grep("^and i'd eat", df_notail[,1]),],
      df_notail[grep("^and i'd die", df_notail[,1]),])
```

Question 2:
Guy at my table's wife got up to go to the bathroom and I asked about dessert and he started telling me about his

```{r}
rbind(df[grep("^about his horticultural", df[,1]),],
      df[grep("^about his spiritual", df[,1]),],
      df[grep("^about his marital", df[,1]),],
      df[grep("^about his financial", df[,1]),])
```

```{r}
rbind(df_notail[grep("^about his horticultural", df_notail[,1]),],
      df_notail[grep("^about his spiritual", df_notail[,1]),],
      df_notail[grep("^about his marital", df_notail[,1]),],
      df_notail[grep("^about his financial", df_notail[,1]),])
```

Question 3:
I'd give anything to see arctic monkeys this

```{r}
head(df[grep("^monkeys this", df[,1]),], 10)
```

```{r}
head(df[grep("^arctic monkeys", df[,1]),], 30)
```

```{r}
rbind(head(df[grep("this weekend", df[,1]),], 1),
      head(df[grep("this month", df[,1]),], 1),
      head(df[grep("this decade", df[,1]),], 1),
      head(df[grep("this morning", df[,1]),], 1))
```

```{r}
rbind(head(df_notail[grep("this weekend", df_notail[,1]),], 1),
      head(df_notail[grep("this month", df_notail[,1]),], 1),
      head(df_notail[grep("this decade", df_notail[,1]),], 1),
      head(df_notail[grep("this morning", df_notail[,1]),], 1))
```

Question 4:
Talking to your mom has the same effect as a hug and helps reduce your

```{r}
head(df[grep("^reduce your", df[,1]),], 5)
```

```{r}
rbind(head(df[grep("^reduce your hunger", df[,1]),], 1),
      head(df[grep("^reduce your stress", df[,1]),], 1),
      head(df[grep("^reduce your happiness", df[,1]),], 1),
      head(df[grep("^reduce your sleepiness", df[,1]),], 1))
```

```{r}
rbind(head(df_notail[grep("^reduce your hunger", df_notail[,1]),], 1),
      head(df_notail[grep("^reduce your stress", df_notail[,1]),], 1),
      head(df_notail[grep("^reduce your happiness", df_notail[,1]),], 1),
      head(df_notail[grep("^reduce your sleepiness", df_notail[,1]),], 1))
```
Answer: stress

Question 5:
When you were in Holland you were like 1 inch away from me but you hadn't time to take a

```{r}
head(df[grep("^take a ", df[,1]),], 5)
```

```{r}
head(df_notail[grep("^take a ", df_notail[,1]),], 5)
```

Question 6:
I'd just like all of these questions answered, a presentation of evidence, and a jury to settle the

```{r}
head(df[grep("^settle the ", df[,1]),], 5)
```

```{r}
head(df_notail[grep("^settle the ", df_notail[,1]),], 5)
```


Question 7:
I can't deal with unsymetrical things. I can't even hold an uneven number of bags of groceries in each

```{r}
rbind(head(df[grep("^in each finger", df[,1]),], 1),
      head(df[grep("^in each arm", df[,1]),], 1),
      head(df[grep("^in each toe", df[,1]),], 1),
      head(df[grep("^in each hand", df[,1]),], 1))
```

```{r}
rbind(head(df_notail[grep("^in each finger", df_notail[,1]),], 1),
      head(df_notail[grep("^in each arm", df_notail[,1]),], 1),
      head(df_notail[grep("^in each toe", df_notail[,1]),], 1),
      head(df_notail[grep("^in each hand", df_notail[,1]),], 1))
```

Answer: hand

Quesiton 8:
Every inch of you is perfect from the bottom to the

```{r}
rbind(head(df[grep("^to the side", df[,1]),], 1),
      head(df[grep("^to the center", df[,1]),], 1),
      head(df[grep("^to the top", df[,1]),], 1),
      head(df[grep("^to the middle", df[,1]),], 1))
```

```{r}
rbind(head(df_notail[grep("^to the side", df_notail[,1]),], 1),
      head(df_notail[grep("^to the center", df_notail[,1]),], 1),
      head(df_notail[grep("^to the top", df_notail[,1]),], 1),
      head(df_notail[grep("^to the middle", df_notail[,1]),], 1))
```
Answer: top

Question 9:
I’m thankful my childhood was filled with imagination and bruises from playing

```{r}
head(df[grep("^from playing", df[,1]),], 30)
```

```{r}
head(df_notail[grep("^from playing", df_notail[,1]),], 10)
```
Answer: outside

Question 10:
I like how the same people are in almost all of Adam Sandler's

```{r}
head(df[grep("^adam sandler's ", df[,1]),], 10)
```

```{r}
df[grep("adam sandler's", df[,1]),]
```
Answer: movies