---
title: "Untitled"
output: html_document
---

# Question 1

The American Community Survey distributes downloadable data about United States communities. Download the 2006 microdata survey about housing for the state of Idaho using download.file() from here:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv

and load the data into R. The code book, describing the variable names is here:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FPUMSDataDict06.pdf



```{r}
library(readr)
library(dplyr)

url <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv"
destfile <- "~/documents/github/datasciencecoursera/3-Getting-And-Cleaning-Data/Quiz/ss06hid.csv"

dataset <- download.file(url, destfile, method = "curl")


ss06hid <- read_csv("Documents/GitHub/datasciencecoursera/3-Getting-And-Cleaning-Data/Quiz/ss06hid.csv")

head(ss06hid)
```

How many properties are worth $1,000,000 or more?

```{r}
count(filter(ss06hid, VAL == 24))
```


# Question 2


Use the data you loaded from Question 1. Consider the variable FES in the code book. Which of the "tidy data" principles does this variable violate?

```{r}
head(ss06hid)
```

- Each tidy data table contains information about only one type of observation.


- Each variable in a tidy data set has been transformed to be interpretable.


**- Tidy data has one variable per column.**


- Numeric values in tidy data can not represent categories.

# Question 3

Download the Excel spreadsheet on Natural Gas Aquisition Program here:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FDATA.gov_NGAP.xlsx



```{r}
url <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FDATA.gov_NGAP.xlsx"
destfile <- "~/documents/github/datasciencecoursera/3-Getting-And-Cleaning-Data/Quiz/d396qusza40orc"
download.file(url, destfile)
```
Read rows 18-23 and columns 7-15 into R and assign the result to a variable called:dat

```{r}
library(xlsx)

rowIndex <- 18:23
colIndx <- 7:15

dat <- read.xlsx(file="Documents/GitHub/datasciencecoursera/3-Getting-And-Cleaning-Data/Quiz/d396qusza40orc", sheetIndex=1,colIndex=colIndx,startRow=18, endRow=23, header=TRUE)

sum(dat$Zip*dat$Ext,na.rm=T)
# 36534720
```

# Question 4

Read the XML data on Baltimore restaurants from here:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Frestaurants.xml

How many restaurants have zipcode 21231?


```{r}
#install.packages("XML")

library(XML)

doc <-  xmlTreeParse("http://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Frestaurants.xml",useInternal = TRUE)

rootnode <- xmlRoot(doc)

sum(xpathSApply(rootnode,"//zipcode",xmlValue) == 21231)
```

# Question 5

The American Community Survey distributes downloadable data about United States communities. Download the 2006 microdata survey about housing for the state of Idaho using download.file() from here:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv

using the fread() command load the data into an R object: DT

The following are ways to calculate the average value of the variable pwgtp15
broken down by sex. 
Using the data.table package, which will deliver the fastest user time?

```{r}
#install.packages("data.table")
library(data.table)
download.file('https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv','./Q5.csv','curl')
print(date())
```


```{r}
DT = fread('./Q5.csv')
system.time(tapply(DT$pwgtp15,DT$SEX,mean))
```

```{r}
system.time(DT[,mean(pwgtp15),by=SEX])
```

```{r}
system.time(mean(DT$pwgtp15,by=DT$SEX))
```

```{r}
system.time(sapply(split(DT$pwgtp15,DT$SEX),mean))
```