---
title: "Analysis of income by college major"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Dependencies

First we load the dataset from the collegeIncome package. Next we load the broom package for tidier
display of regression output and the dplyr package for working with data frames.

```{r message=FALSE, warning=FALSE}
library(collegeIncome)
data(college)
library(broom)
library(dplyr)
```

# Exploratory data analysis

Let’s take a look at the data to get acquainted with the contents.

```{r}
head(college)

summary(college)
```

What are the different categories of college majors?

```{r}
table(college$major_category)
```

Only one major falls into the “Interdisciplinary” category. We certainly cannot estimate an effect for this
category, so we will remove it.

```{r}
college <- college %>%
filter(major_category != "Interdisciplinary")
table(college$major_category)
```

Median income in a category is a useful measure because it indicates an income level that is typical in that
category. Later we will used median income as our outcome measure in linear regression, but it is useful to
look at plots of how the three different income measures (25th, 50th, and 75th percentile of income among
those reporting income in the survey) vary across categories.

```{r}
par(mar = c(13,4.5,2,0.5))
boxplot(p75th ~ major_category, data = college, main = "75th percentile", las = 2, cex.axis = 0.75)
```

25th percentile

```{r}
boxplot(p25th ~ major_category, data = college, main = "25th percentile", las = 2, cex.axis = 0.75)
```

50th percentile

```{r}
boxplot(median ~ major_category, data = college, main = "50th percentile", las = 2, cex.axis = 0.50)
```

There doesn’t seem to be considerable variation across categories for either of these three measures, so we
will stick with using the median as our outcome measure.

# Linear regression

Because the income information is defined to pertain to full-time, year-round workers, we will look at other
characteristics: namely, gender effects and effects related to type of job (jobs requiring a college degree and
jobs that are low-wage service positions). We can fit this linear model and view inference results with the
following commands:

```{r}
lmfit <- lm(median ~ major_category+perc_women+perc_college_jobs+perc_low_wage_jobs, data = college)
summary(lmfit)
```

Holding constant gender distributions and skill category distributions, we don’t see much effect of major
category (reference category is Agriculture & Natural Resources - how can we tell?). In particular, it is
important to consider the multiple hypothesis testing issue here as we have many different major categories.
Considering this, we really don’t see much of a category effect on income. Further the F-statistic results at
the bottom do not suggest that the variables included have an impact on median income.

We can check if regression assumptions are met with diagnostic plotting:

```{r}
resid <- residuals(lmfit)
fitted <- fitted.values(lmfit)
plot(density(resid), xlab = "Residuals", ylab = "Density", main = "Residual distribution")
```

```{r}
plot(fitted, resid, xlab = "Predicted values", ylab = "Residuals")
abline(h = 0, col = "red", lty = "dashed")
```

Normality assumptions don’t seem far off (approximately), and heteroskedasticity doesn’t seem to be an issue.
Perhaps there is one outlier, but that is unlikely to have changed the overall results too much.
Overall there doesn’t seem to be an effect of college major category on median income in this study.

FYI: A slightly neater display of the inference results for the coefficients can be obtained with the tidy
function in the broom package:

```{r}
tidy(lmfit)
```
