---
title: "Plotly Assignment"
author: "Petra Mala"
date: "20/10/2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.table.format = "html") 
```

# Simple plotly example - iris dataset



Settings
----
We are going to plot 'Petal Length' vs 'Petal Width', by 'Species'
```{r message=FALSE, warning=FALSE}
library(plotly)
library(datasets)
library(dplyr)
data(iris)
```

Formatting
----
```{r echo=TRUE, message=FALSE, warning=FALSE}
f <- list(
  family = "Courier New, monospace",
  size = 11,   color = "#6e6e6e" )
x <- list(
  title = "Petal Length",   titlefont = f )
y <- list(
  title = "Petal Width",   titlefont = f )
```


Plotting
----
```{r message=FALSE, warning=FALSE, echo=FALSE}
plot_ly(iris, x = ~Petal.Length, y = ~Petal.Width,  type="scatter", mode = "markers" , color = ~Species , 
        colors = "Set1",  
        marker=list( size=20 , opacity=0.5) )%>%
        layout(xaxis=x, yaxis=y) 
```

Interpretation
----

The iris dataset in R provides measurements (in cm) of sepal length and width, and petal length and width of the following 3 species of iris: Iris setosa, versicolor, and virginica. This plot shows relaitonship between petal length and petal width. Different species are shown in different colors.

  

References
----


- "Presentations with ioslides - Overview" 
<http://rmarkdown.rstudio.com/ioslides_presentation_format.html>

- "Iris Dataset" 
<https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/iris.html>