---
title: "Shiny Application and Reproducible Pitch"
author: "Petra Mala"
date: "24/11/2019"
output: slidy_presentation
---

## Project
This peer assessed assignment has two parts. First we will create a Shiny Application and deploy it on Rstudio server.
Secondly, we will use Slidify or Rstudio Presenter to prepare a reproducible pitch presentation about your application.

## Objectives

- Developing simple application using R Studio

UI:

- Define your distribution type
- Select sample size
- If Distribution Type is Normal then allow user to select mean and    standard deviation
- If Distribution Type is Exponential then allow user to select        lambda



## UI Configuration

ui.R

```{r eval=F, echo=T}
shinyServer(
  pageWithSidebar(
    headerPanel("My First Shiny App"),
    sidebarPanel(
      selectInput("Distribution", "Please Select Distribution Type",
                  choices=c("Normal", "Exponential")),
      sliderInput("sampleSize", "Please Select Sample Size: ",
                  min=100, max=5000, value= 1000, step=100 ),
      conditionalPanel(
        condition = "input.Distribution == 'Normal'",
        textInput("mean","Please Select the mean",10),
        textInput("sd","Please Select Standard Deviation", 3)
      ),
      conditionalPanel(
        condition = "input.Distribution == 'Exponential'",
        textInput("lambda", "Please Select Exponential Lambda:",1)
      )
    ),
    mainPanel(plotOutput("myPlot"))
  )
)
```


## Server Configuration

server.R

```{r eval=F, echo=T}
shinyServer(
  
  function(input, output, sessionInfo) {
    
    output$myPlot <- renderPlot({
      
      distType <- input$Distribution
      size <- input$sampleSize
      
      if(distType == "Normal") {
        randomVec <- rnorm(size, mean = as.numeric(input$mean))
      } else {
        randomVec <- rnorm(size, mean = as.numeric(input$mean), sd=as.numeric(input$sd))
      }
      hist(randomVec, col="blue")
    })
  }
)
```

## ShinyApp Code

Git repository at: https://github.com/MalaPetra/DataScienceCourse/tree/master/ShinyApp