---
title: "Quizzes"
output: html_document
---
# Quiz 1

### Question 1

Which of the following are components in building a machine learning algorithm?

=> creating features

### Question 2

Suppose we build a prediction algorithm on a data set and it is 100% accurate on that data set. Why might the algorithm not work well if we collect a new data set?

=> Our algorithm may be overfitting the training data, predicting both the signal and the noice

### Question 3

What are typical sizes for the training and test sets?

=> 80% training set, 20% test set

### Question 4

What are some common error rates for predicting binary variables (i.e. variables with two possible values like yes/no, disease/normal, clicked/didn't click)? Check the correct answer(s).

=> Predictive value of the positive


### Question 5

Suppose that we have created a machine learning algorithm that predicts whether a link will be clicked with 99% sensitivity and 99% specificity. The rate the link is clicked is 1/1000 of visits to a website. If we predict the link will be clicked on a specific visit, what is the probability it will actually be clicked?

=> 9%


# Quiz 2

### Question 1
Load the Alzheimer's disease data using the commands:


```{r}
#install.packages("AppliedPredictiveModeling")
library(AppliedPredictiveModeling)
data(AlzheimerDisease)
install.packages("caret")
library(caret)
```

Which of the following commands will create non-overlapping training and test sets with about 50% of the observations assigned to each?

```{r}
adData = data.frame(diagnosis,predictors)  
testIndex = createDataPartition(diagnosis, p = 0.50,list=FALSE)  
training = adData[-testIndex,]  
testing = adData[testIndex,]  
```

### Question 2
Load the cement data using the commands:

```{r}
library(AppliedPredictiveModeling)  
data(concrete)  
library(caret)  
set.seed(1000)  
inTrain = createDataPartition(mixtures$CompressiveStrength, p = 3/4)[[1]]  
training = mixtures[ inTrain,]  
testing = mixtures[-inTrain,]
```

Make a plot of the outcome (CompressiveStrength) versus the index of the samples. Color by each of the variables in the data set (you may find the cut2() function in the Hmisc package useful for turning continuous covariates into factors). What do you notice in these plots?

```{r}
library(GGally)
library(Hmisc)
## Using ggpair
training2 <- training
#cut CompressiveStrength into 4 levels.  This is the only way to work with colour in ggpair
training2$CompressiveStrength <- cut2(training2$CompressiveStrength, g=4)
ggpairs(data = training2, columns = c("FlyAsh","Age","CompressiveStrength"), mapping = ggplot2::aes(colour = CompressiveStrength))
```

Make a plot of the outcome (CompressiveStrength) versus the index of the samples. Color by each of the variables in the data set (you may find the cut2() function in the Hmisc package useful for turning continuous covariates into factors). What do you notice in these plots?

Before we need to get the names of the columns to subset

```{r}
names <- colnames(concrete)
names <- names[-length(names)]
```

Now let’s make a quick feature plot to see if there is any relation between the outcome CompressiveStrength and the rest of the parameters in the data:

```{r}
featurePlot(x = training[, names], y = training$CompressiveStrength, plot = "pairs")
```

We can observe on this plot that there is no relation between the outcome and any of the other variables in data set

Now we will make a plot of the outcome as a function of the index

```{r}
index <- seq_along(1:nrow(training))
ggplot(data = training, aes(x = index, y = CompressiveStrength)) + geom_point() + theme_bw()
```

Now we see this figure that there is a step-like pattern in the data that could be explained by one or more variable in the data. From this plot we should probably cut the outcome in 4 categories.

```{r}
cutCS <- cut2(training$CompressiveStrength, g = 4)
summary(cutCS)
```

Then, we make a plot of the categorized outcome outcome

```{r}
ggplot(data = training, aes(y = index, x = cutCS)) + geom_boxplot() + geom_jitter(col = "blue") + theme_bw()
```


Now the step is better seen in the above plot. As we can see this plot the step-like pattern is more clear now.

Now we’ll make a plot of the categorized income as function of the rest of the variables

```{r}
featurePlot(x = training[, names], y = cutCS, plot = "box")
```

Once more, none of the variables in the data can explaing the step-like behaviour in the outcome.

There is a non-random pattern in the plot of the outcome versus index that does not appear to be perfectly explained by any predictor suggesting a variable may be missing.

### Question 3

```{r}
#Load the cement data using the commands:  
library(AppliedPredictiveModeling)  
data(concrete)  
library(caret)  
set.seed(1000)  
inTrain = createDataPartition(mixtures$CompressiveStrength, p = 3/4)[[1]]  
training = mixtures[ inTrain,]  
testing = mixtures[-inTrain,]  
```

Make a histogram and confirm the SuperPlasticizer variable is skewed. Normally you might use the log transform to try to make the data more symmetric. Why would that be a poor choice for this variable?

```{r}
par(mfrow = c(2,1))
hist(training$Superplasticizer)
hist(log(training$Superplasticizer + 1))
```

```{r}
ggplot(data = training, aes(x = Superplasticizer)) + geom_histogram() + theme_bw()
```

There are values of zero so when you take the log() transform those values will be -Inf.



### Question 4

```{r}
library(caret)  
library(AppliedPredictiveModeling)  
set.seed(3433)
data(AlzheimerDisease)  
adData = data.frame(diagnosis,predictors)  
inTrain = createDataPartition(adData$diagnosis, p = 3/4)[[1]]
training = adData[ inTrain,]  
testing = adData[-inTrain,]  
```

Find all the predictor variables in the training set that begin with IL. Perform principal components on these variables with the preProcess() function from the caret package. Calculate the number of principal components needed to capture 90% of the variance. How many are there?

```{r}
trainingIL <- training[,grep("^IL", names(training))]
procTrain <- preProcess(trainingIL, method = "pca", thresh = 0.9 )
procTrain
```

### Question 5

```{r}
library(caret)
library(AppliedPredictiveModeling)
set.seed(3433)
data(AlzheimerDisease)
adData = data.frame(diagnosis,predictors)
inTrain = createDataPartition(adData$diagnosis, p = 3/4)[[1]]
training = adData[ inTrain,]
testing = adData[-inTrain,]
```

Create a training data set consisting of only the predictors with variable names beginning with IL and the diagnosis. Build two predictive models, one using the predictors as they are and one using PCA with principal components explaining 80% of the variance in the predictors. Use method="glm" in the train function.

What is the accuracy of each method in the test set? Which is more accurate?

```{r}

# grep all columns with IL and diagnosis in the traning and testing set
trainingIL <- training[,grep("^IL|diagnosis", names(training))]
testingIL <- testing[,grep("^IL|diagnosis", names(testing))]

# non-PCA
model <- train(diagnosis ~ ., data = trainingIL, method = "glm")
predict_model <- predict(model, newdata= testingIL)
matrix_model <- confusionMatrix(predict_model, testingIL$diagnosis)
matrix_model$overall[1]
```

```{r}
# PCA
modelPCA <- train(diagnosis ~., data = trainingIL, method = "glm", preProcess = "pca",trControl=trainControl(preProcOptions=list(thresh=0.8)))
matrix_modelPCA <- confusionMatrix(testingIL$diagnosis, predict(modelPCA, testingIL))
matrix_modelPCA$overall[1]
```

# Quiz 3

### Question 1


For this quiz we will be using several R packages. R package versions change over time, the right answers have been checked using the following versions of the packages.

- AppliedPredictiveModeling: v1.1.6
- caret: v6.0.47
- ElemStatLearn: v2012.04-0
- pgmm: v1.1
- rpart: v4.1.8

If you aren’t using these versions of the packages, your answers may not exactly match the right answer, but hopefully should be close.

Load the cell segmentation data from the AppliedPredictiveModeling package using the commands:

```{r}
library(AppliedPredictiveModeling)
data(segmentationOriginal)
library(caret)
```

1. Subset the data to a training set and testing set based on the Case variable in the data set.

2. Set the seed to 125 and fit a CART model with the rpart method using all predictor variables and default caret settings.

3. In the final model what would be the final model prediction for cases with the following variable values:

- TotalIntench2 = 23,000; FiberWidthCh1 = 10; PerimStatusCh1=2

- TotalIntench2 = 50,000; FiberWidthCh1 = 10;VarIntenCh4 = 100

- TotalIntench2 = 57,000; FiberWidthCh1 = 8;VarIntenCh4 = 100

- FiberWidthCh1 = 8;VarIntenCh4 = 100; PerimStatusCh1=2

```{r}
library(rattle)
str(segmentationOriginal)

trainSet <- segmentationOriginal[segmentationOriginal$Case =="Train",]
testSet <- segmentationOriginal[segmentationOriginal$Case =="Test",]

set.seed(125)
model_rpart <- train(Class~.,data=trainSet,method="rpart")
```

```{r}
fancyRpartPlot(model_rpart$finalModel)
```

Answer:  a: PS;b: WS;c: PS;d: Not possible to predict.


### Question 2

If K is small in a K-fold cross validation is the bias in the estimate of out-of-sample (test set) accuracy smaller or bigger? If K is small is the variance in the estimate of out-of-sample (test set) accuracy smaller or bigger. Is K large or small in leave one out cross validation?

Answer: The bias is larger and the variance is smaller. Under leave one out cross validation K is equal to the sample size.

### Question 3

```{r}
library(pgmm)
data(olive)
olive = olive[,-1]
```

These data contain information on 572 different Italian olive oils from multiple regions in Italy. Fit a classification tree where Area is the outcome variable. Then predict the value of area for the following data frame using the tree command with all defaults

```{r}
newdata = as.data.frame(t(colMeans(olive)))
```

What is the resulting prediction? Is the resulting prediction strange? Why or why not?

```{r}
str(olive)
```

```{r}
table(olive$Area)
```

```{r}
olive_rpart <- train(Area~.,data=olive,method="rpart")
```

```{r}
fancyRpartPlot(olive_rpart$finalModel)
```

```{r}
predict(olive_rpart,newdata=newdata)
```

Answer: 2.783. It is strange because Area should be a qualitative variable - but tree is reporting the average value of Area as a numeric variable in the leaf predicted for newdata.

### Question 4

Load the South Africa Heart Disease Data and create training and test sets with the following code:

```{r}
library(SAheart)
data(SAheart)
set.seed(8484)
train = sample(1:dim(SAheart)[1],size=dim(SAheart)[1]/2,replace=F)
trainSA = SAheart[train,]
testSA = SAheart[-train,]
```

Then set the seed to 13234 and fit a logistic regression model (method="glm", be sure to specify family="binomial") with Coronary Heart Disease (chd) as the outcome and age at onset, current alcohol consumption, obesity levels, cumulative tabacco, type-A behavior, and low density lipoprotein cholesterol as predictors. Calculate the misclassification rate for your model using this function and a prediction on the "response" scale:

```{r}
missClass = function(values,prediction){sum(((prediction > 0.5)*1) != values)/length(values)}
```

```{r}
set.seed(13234)

# definition of the training model
regModel <- train(chd~age+alcohol+obesity+tobacco+typea+ldl,data=trainSA,method="glm",family="binomial")
```

```{r}
# computation of the misclasssification on the training set and test set
missClassTrain <- missClass(trainSA$chd,predict(regModel,newdata=trainSA))
missClassTest <- missClass(testSA$chd,predict(regModel,newdata=testSA))
missClassTrain
```

```{r}
missClassTest
```

Test Set Misclassification: 0.31
Training Set: 0.27.
  
### Question 5

Load the vowel.train and vowel.test data sets:

```{r}
library(ElemStatLearn)
data(vowel.train)
data(vowel.test)
```

Set the variable y to be a factor variable in both the training and test set. Then set the seed to 33833. Fit a random forest predictor relating the factor variable y to the remaining variables. Read about variable importance in random forests here: http://www.stat.berkeley.edu/~breiman/RandomForests/cc_home.htm#ooberr.

The caret package uses by default the Gini importance.

Calculate the variable importance using the varImp function in the caret package. What is the order of variable importance?

[NOTE: Use randomForest() specifically, not caret, as there’s been some issues reported with that approach. 11/6/2016]

```{r}
set.seed(33833)
str(vowel.train)
```

```{r}
library(randomForest)
```

```{r}
vowel.train$y <- as.factor(vowel.train$y)
vowel.test$y <- as.factor(vowel.test$y)
modelRF <- randomForest(y~.,data=vowel.train)
order(varImp(modelRF),decreasing=TRUE)
```

The order of the variables is:
x.2, x.1, x.5, x.6, x.8, x.4, x.9, x.3, x.7,x.10.

### Quiz 3 Notes

from quiz 3: https://datascience.study.foletta.org/C8_Practical_Machine_Learning/w3_notes.html#measures-of-impurity

# Quiz 4


### Question 1

For this quiz we will be using several R packages. R package versions change over time, the right answers have been checked using the following versions of the packages.

- AppliedPredictiveModeling: v1.1.6
- caret: v6.0.47
- ElemStatLearn: v2012.04-0
- pgmm: v1.1
- rpart: v4.1.8
- gbm: v2.1
- lubridate: v1.3.3
- forecast: v5.6
- e1071: v1.6.4

If you aren’t using these versions of the packages, your answers may not exactly match the right answer, but hopefully should be close.

Load the vowel.train and vowel.test data sets:

```{r}
library(ElemStatLearn)
data(vowel.train)
data(vowel.test)
```

Set the variable y to be a factor variable in both the training and test set. Then set the seed to 33833. Fit (1) a random forest predictor relating the factor variable y to the remaining variables and (2) a boosted predictor using the “gbm” method. Fit these both with the train() command in the caret package.

What are the accuracies for the two approaches on the test data set? What is the accuracy among the test set samples where the two methods agree?

```{r}
vowel.train$y <- as.factor(vowel.train$y)
vowel.test$y <- as.factor(vowel.test$y)
set.seed(33833)
modRF <- train(y ~ ., data=vowel.train, method="rf") #, trControl=trainControl("cv"), number=3)
modBoost <- train(y ~ ., data=vowel.train, method="gbm", verbose=FALSE)

predRF <- predict(modRF, vowel.test)
predBoost <- predict(modBoost, vowel.test)
agreedIndex <- predRF == predBoost

cfmRf <- confusionMatrix(vowel.test$y, predRF)
cfmBoost <- confusionMatrix(vowel.test$y, predBoost)
cfmAgreed <- confusionMatrix(vowel.test$y[agreedIndex], predRF[agreedIndex])

cfmRf$overall["Accuracy"]

##  Accuracy 
## 0.6147186
```

```{r}
cfmBoost$overall["Accuracy"]
##  Accuracy 
## 0.5367965
```

```{r}
cfmAgreed$overall["Accuracy"]
##  Accuracy 
## 0.6656051
```

### Question 2

Load the Alzheimer’s data using the following commands

```{r}
library(gbm)
set.seed(3433)
library(AppliedPredictiveModeling)
data(AlzheimerDisease)
adData = data.frame(diagnosis,predictors)
inTrain = createDataPartition(adData$diagnosis, p = 3/4)[[1]]
training = adData[ inTrain,]
testing = adData[-inTrain,]
```

Set the seed to 62433 and predict diagnosis with all the other variables using a random forest (“rf”), boosted trees (“gbm”) and linear discriminant analysis (“lda”) model. Stack the predictions together using random forests (“rf”). What is the resulting accuracy on the test set? Is it better or worse than each of the individual predictions?

```{r}
set.seed(62433)
modRF2 <- train(diagnosis ~ ., data=training, method="rf") #, trControl=trainControl("cv"), number=3)
modBoost2 <- train(diagnosis ~ ., data=training, method="gbm", verbose=FALSE)
modLDA2 <- train(diagnosis ~ ., data=training, method="lda", verbose=FALSE)

predRF2 <- predict(modRF2, testing)
predBoost2 <- predict(modBoost2, testing)
predLDA2 <- predict(modLDA2, testing)

dataCombined <- data.frame(predRF2, predBoost2, predLDA2, diagnosis=testing$diagnosis)
modCombined <- train(diagnosis ~ ., data=dataCombined, method="rf", verbose=FALSE)
## note: only 2 unique complexity parameters in default grid. Truncating the grid to 2 .
```
```{r}
predCombined <- predict(modCombined, dataCombined)

cfmRF2 <- confusionMatrix(testing$diagnosis, predRF2)
cfmBoost2 <- confusionMatrix(testing$diagnosis, predBoost2)
cfmLDA2 <- confusionMatrix(testing$diagnosis, predLDA2)
cfmCombined <- confusionMatrix(testing$diagnosis, predCombined)


cfmRF2$overall["Accuracy"]
##  Accuracy 
## 0.7682927
```

```{r}
cfmBoost2$overall["Accuracy"]
##  Accuracy 
## 0.7926829
```

```{r}
cfmLDA2$overall["Accuracy"]
##  Accuracy 
## 0.7682927
```

```{r}
cfmCombined$overall["Accuracy"]
## Accuracy 
## 0.804878
```

## Question 3

Load the concrete data with the commands:

```{r}
set.seed(3523)

library(AppliedPredictiveModeling)

data(concrete)

inTrain = createDataPartition(concrete$CompressiveStrength, p = 3/4)[[1]]

training = concrete[ inTrain,]

testing = concrete[-inTrain,]
```

Set the seed to 233 and fit a lasso model to predict Compressive Strength. Which variable is the last coefficient to be set to zero as the penalty increases? (Hint: it may be useful to look up ?plot.enet).

```{r}
set.seed(233) 
modLasso <- train(CompressiveStrength ~ ., data=training, method="lasso") plot.enet(modLasso$finalModel,  xvar="penalty", use.color=TRUE)
```

### Question 4
Load the data on the number of visitors to the instructors blog from here:

https://d396qusza40orc.cloudfront.net/predmachlearn/gaData.csv

```{r}
library(lubridate) # For year() function below
dat = read.csv("gaData.csv")
training = dat[year(dat$date) < 2012,]
testing = dat[(year(dat$date)) > 2011,]
tstrain = ts(training$visitsTumblr)
```

Fit a model using the bats() function in the forecast package to the training time series. Then forecast this model for the remaining time points. For how many of the testing points is the true value within the 95% prediction interval bounds?

```{r}
modBats <- bats(tstrain)
forecastObj <- forecast(modBats, level=95, h=nrow(testing))
betweenVal <- sum(testing$visitsTumblr > forecastObj$lower &  testing$visitsTumblr < forecastObj$upper)
betweenVal / nrow(testing) * 100
## [1] 96.17021
```

### Question 5
Load the concrete data with the commands:

```{r}
set.seed(3523)
library(AppliedPredictiveModeling)
data(concrete)
inTrain = createDataPartition(concrete$CompressiveStrength, p = 3/4)[[1]]
training = concrete[ inTrain,]
testing = concrete[-inTrain,]
```

Set the seed to 325 and 􀃒t a support vector machine using the e1071 package to predict Compressive Strength using the default settings. Predict on the testing set. What is the RMSE?

```{r}
set.seed(325)
modSvm <- svm(CompressiveStrength ~ ., data = training)
predSvm <- predict(modSvm, testing)
accSvm <- accuracy(predSvm, testing$CompressiveStrength)
data.frame(accSvm)["RMSE"]
##              RMSE
## Test set 6.715009
```